Water, High-Altitude Condensates, and \added{Possible} Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}

### 2018-02-26 ApJL_submission/ms.tex
final

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. \added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to methane depletion.} \deleted{and we infer a water abundance consistent with expectations for a solar abundance pattern. On the other hand, methane is depleted relative to expectations (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.}  The spectral features are smaller than predicted for a cloudless atmosphere, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. \added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) produce clouds at this pressure level.} \deleted{however, we find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.}  Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

\begin{deluxetable*}{LLLLLLL}
	\tablecolumns{7}
	\tablewidth{0pt}
	\tablecaption{WASP-107b Transmission Spectrum and Limb Darkening Coefficients\label{tab:tspec}}
	\tablehead{
		\colhead{Wavelength} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{$u_1$} & \colhead{$u_2$} & \colhead{$u$} \\
	\colhead{(micron)} & \colhead{(Method A)\tablenotemark{*}} & \colhead{(Method B)} & \colhead{(Method C)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}
		\startdata
		1.121 - 1.145 & 0.020641 \pm 5.9e-05 & 0.020674 \pm 7.8e-05 & 0.020521 \pm 7.4e-05 & 0.38 & 0.13 & 0.44 \pm 0.011 \\
		1.145 - 1.171 & 0.020733 \pm 5.5e-05 & 0.020825 \pm 7.5e-05 & 0.020575 \pm 7.7e-05 & 0.37 & 0.14 & 0.41 \pm 0.012 \\
		1.171 - 1.196 & 0.020505 \pm 5.6e-05 & 0.020655 \pm 7.4e-05 & 0.020388 \pm 8.2e-05 & 0.36 & 0.14 & 0.4 \pm 0.012 \\
		1.196 - 1.222 & 0.020498 \pm 5.4e-05 & 0.020580 \pm 7.4e-05 & 0.020345 \pm 6.6e-05 & 0.36 & 0.15 & 0.41 \pm 0.012 \\
		1.222 - 1.248 & 0.020455 \pm 5.9e-05 & 0.020620 \pm 7.8e-05 & 0.020327 \pm 7.7e-05 & 0.36 & 0.15 & 0.39 \pm 0.012 \\
		1.248 - 1.272 & 0.020492 \pm 5.0e-05 & 0.020649 \pm 6.3e-05 & 0.020360 \pm 7.6e-05 & 0.35 & 0.16 & 0.39 \pm 0.01 \\
		1.272 - 1.298 & 0.020620 \pm 6.2e-05 & 0.020702 \pm 9.0e-05 & 0.020475 \pm 8.3e-05 & 0.34 & 0.17 & 0.42 \pm 0.014 \\
		1.298 - 1.323 & 0.020739 \pm 5.0e-05 & 0.020883 \pm 6.7e-05 & 0.020602 \pm 8.6e-05 & 0.34 & 0.17 & 0.39 \pm 0.011 \\
		1.323 - 1.349 & 0.020660 \pm 5.7e-05 & 0.020789 \pm 8.7e-05 & 0.020489 \pm 8.3e-05 & 0.33 & 0.18 & 0.4 \pm 0.014 \\
		1.349 - 1.374 & 0.020858 \pm 4.8e-05 & 0.020916 \pm 7.6e-05 & 0.020706 \pm 7.9e-05 & 0.32 & 0.19 & 0.42 \pm 0.012 \\
		1.374 - 1.401 & 0.020794 \pm 4.8e-05 & 0.020864 \pm 8.1e-05 & 0.020653 \pm 7.9e-05 & 0.31 & 0.20 & 0.41 \pm 0.012 \\
		1.401 - 1.425 & 0.020888 \pm 5.2e-05 & 0.020961 \pm 6.1e-05 & 0.020776 \pm 7.0e-05 & 0.30 & 0.21 & 0.41 \pm 0.01 \\
		1.425 - 1.452 & 0.020821 \pm 6.2e-05 & 0.020933 \pm 7.8e-05 & 0.020708 \pm 8.5e-05 & 0.29 & 0.21 & 0.39 \pm 0.011 \\
		1.452 - 1.476 & 0.020691 \pm 5.1e-05 & 0.020876 \pm 7.8e-05 & 0.020556 \pm 8.7e-05 & 0.28 & 0.22 & 0.36 \pm 0.012 \\
		1.476 - 1.502 & 0.020682 \pm 6.9e-05 & 0.020890 \pm 8.7e-05 & 0.020549 \pm 1.0e-04 & 0.26 & 0.23 & 0.34 \pm 0.014 \\
		1.502 - 1.528 & 0.020679 \pm 6.7e-05 & 0.020763 \pm 9.6e-05 & 0.020523 \pm 8.8e-05 & 0.25 & 0.23 & 0.37 \pm 0.015 \\
		1.528 - 1.552 & 0.020509 \pm 6.0e-05 & 0.020667 \pm 9.3e-05 & 0.020359 \pm 1.0e-04 & 0.23 & 0.25 & 0.33 \pm 0.014 \\
		1.552 - 1.579 & 0.020480 \pm 6.4e-05 & 0.020551 \pm 1.2e-04 & 0.020376 \pm 1.3e-04 & 0.22 & 0.24 & 0.38 \pm 0.019 \\
		1.579 - 1.603 & 0.020500 \pm 5.6e-05 & 0.020542 \pm 1.1e-04 & 0.020321 \pm 1.1e-04 & 0.20 & 0.24 & 0.33 \pm 0.019 \\
		1.603 - 1.629 & 0.020514 \pm 6.5e-05 & 0.020755 \pm 1.1e-04 & 0.020392 \pm 1.1e-04 & 0.19 & 0.25 & 0.29 \pm 0.019 \\
		\enddata
		\tablenotetext{*}{Our retrieval analysis uses the transit depths from Method A.}
		\tablecomments{comments}
	\end{deluxetable*}

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit14}.}

@@

We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predicted the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}. \added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%. We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well. More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.}

@@

>>>>

\begin{abstract}
	The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar. The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern. \added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.}  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height. A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations. \added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.} Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

\begin{deluxetable*}{LLLLLLL}
	\tablecolumns{7}
	\tablewidth{0pt}
	\tablecaption{WASP-107b Transmission Spectrum and Limb Darkening Coefficients\label{tab:tspec}}
	\tablehead{
		\colhead{Wavelength} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{$u_1$} & \colhead{$u_2$} & \colhead{$u$} \\
	\colhead{(micron)} & \colhead{(Method A)} & \colhead{(Method B)} & \colhead{(Method C)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}
		\startdata
		1.121 - 1.145 & 0.020641 \pm 5.9e-05 & 0.020674 \pm 7.8e-05 & 0.020521 \pm 7.4e-05 & 0.38 & 0.13 & 0.44 \pm 0.011 \\
		1.145 - 1.171 & 0.020733 \pm 5.5e-05 & 0.020825 \pm 7.5e-05 & 0.020575 \pm 7.7e-05 & 0.37 & 0.14 & 0.41 \pm 0.012 \\
		1.171 - 1.196 & 0.020505 \pm 5.6e-05 & 0.020655 \pm 7.4e-05 & 0.020388 \pm 8.2e-05 & 0.36 & 0.14 & 0.4 \pm 0.012 \\
		1.196 - 1.222 & 0.020498 \pm 5.4e-05 & 0.020580 \pm 7.4e-05 & 0.020345 \pm 6.6e-05 & 0.36 & 0.15 & 0.41 \pm 0.012 \\
		1.222 - 1.248 & 0.020455 \pm 5.9e-05 & 0.020620 \pm 7.8e-05 & 0.020327 \pm 7.7e-05 & 0.36 & 0.15 & 0.39 \pm 0.012 \\
		1.248 - 1.272 & 0.020492 \pm 5.0e-05 & 0.020649 \pm 6.3e-05 & 0.020360 \pm 7.6e-05 & 0.35 & 0.16 & 0.39 \pm 0.01 \\
		1.272 - 1.298 & 0.020620 \pm 6.2e-05 & 0.020702 \pm 9.0e-05 & 0.020475 \pm 8.3e-05 & 0.34 & 0.17 & 0.42 \pm 0.014 \\
		1.298 - 1.323 & 0.020739 \pm 5.0e-05 & 0.020883 \pm 6.7e-05 & 0.020602 \pm 8.6e-05 & 0.34 & 0.17 & 0.39 \pm 0.011 \\
		1.323 - 1.349 & 0.020660 \pm 5.7e-05 & 0.020789 \pm 8.7e-05 & 0.020489 \pm 8.3e-05 & 0.33 & 0.18 & 0.4 \pm 0.014 \\
		1.349 - 1.374 & 0.020858 \pm 4.8e-05 & 0.020916 \pm 7.6e-05 & 0.020706 \pm 7.9e-05 & 0.32 & 0.19 & 0.42 \pm 0.012 \\
		1.374 - 1.401 & 0.020794 \pm 4.8e-05 & 0.020864 \pm 8.1e-05 & 0.020653 \pm 7.9e-05 & 0.31 & 0.20 & 0.41 \pm 0.012 \\
		1.401 - 1.425 & 0.020888 \pm 5.2e-05 & 0.020961 \pm 6.1e-05 & 0.020776 \pm 7.0e-05 & 0.30 & 0.21 & 0.41 \pm 0.01 \\
		1.425 - 1.452 & 0.020821 \pm 6.2e-05 & 0.020933 \pm 7.8e-05 & 0.020708 \pm 8.5e-05 & 0.29 & 0.21 & 0.39 \pm 0.011 \\
		1.452 - 1.476 & 0.020691 \pm 5.1e-05 & 0.020876 \pm 7.8e-05 & 0.020556 \pm 8.7e-05 & 0.28 & 0.22 & 0.36 \pm 0.012 \\
		1.476 - 1.502 & 0.020682 \pm 6.9e-05 & 0.020890 \pm 8.7e-05 & 0.020549 \pm 1.0e-04 & 0.26 & 0.23 & 0.34 \pm 0.014 \\
		1.502 - 1.528 & 0.020679 \pm 6.7e-05 & 0.020763 \pm 9.6e-05 & 0.020523 \pm 8.8e-05 & 0.25 & 0.23 & 0.37 \pm 0.015 \\
		1.528 - 1.552 & 0.020509 \pm 6.0e-05 & 0.020667 \pm 9.3e-05 & 0.020359 \pm 1.0e-04 & 0.23 & 0.25 & 0.33 \pm 0.014 \\
		1.552 - 1.579 & 0.020480 \pm 6.4e-05 & 0.020551 \pm 1.2e-04 & 0.020376 \pm 1.3e-04 & 0.22 & 0.24 & 0.38 \pm 0.019 \\
		1.579 - 1.603 & 0.020500 \pm 5.6e-05 & 0.020542 \pm 1.1e-04 & 0.020321 \pm 1.1e-04 & 0.20 & 0.24 & 0.33 \pm 0.019 \\
		1.603 - 1.629 & 0.020514 \pm 6.5e-05 & 0.020755 \pm 1.1e-04 & 0.020392 \pm 1.1e-04 & 0.19 & 0.25 & 0.29 \pm 0.019 \\
		\enddata
		\tablecomments{Our retrieval analysis uses the transit depths from Method A. Transit depths from Methods A and B use quadratic limb darkening coefficients calculated from a PHOENIX stellar model (listed in columns 5 and 6). Method C fits a linear limb darkening coefficient (fit results listed in column 7).}
	\end{deluxetable*}

@@

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}

@@

We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predicted the abundance of hydrocarbon haze precursors from previously published $50\times$ solar photochemical models for GJ 436b \citep{line11, morley17}. \added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%. We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well. More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.}

@@

\bibitem[{{de Wit} \& {Seager}(2013)}]{dewit13}
{de Wit}, J., \& {Seager}, S. 2013, Science, 342, 1473
